\thispagestyle{empty}

\begin{center}
  {\bf \Huge Acknowledgements}
\end{center}

\vspace{4cm}


\emph{
  In my experience, everything we do in our lives has some kind of contribution from the people that live around us at home, at school, at work, everywhere. Unless you are a hermit, of course. And there are different kinds of support people can give you, depending on the relation you have with them. 
On a more affective side, I’d like to thank my family for always believing in me and in what I do, and supporting me in my choices. My girlfriend and my friends, for their support and the great time I always spend with them and that helps me recovering from the overwhelming amount of commitments of every day life. And my course mates, for all the time we studied together sharing knowledge and useful suggestions.
On the professional side, first I’d like to express my deepest gratitude to my Supervisor and Professor Alberto Montresor for the passion he transmitted me while teaching Algorithms on the second year, for involving me in interesting projects and for his supervision in this last work of my bachelor degree.
I’d also like to thank my Erasmus Professor Erki Eassar for his skill in making me appreciate the course of Databases and for sharing me useful suggestions and materials even months after being his student in Tallinn University of Technology.
At last, I’d like to acknowledge my internship tutors in Tai, Andrea Carpineti and David Votino, for their technical and motivational suggestions about the project from which this dissertation has origin, and my colleague Bruno Graziano for his precious help in configuring and maintaining the system of virtual machines that hosted my clients and my Mongo nodes for the main tests. 
Have a good reading.
}

